---
title: "Print, Summary, and Shiny App"
author: "Stat 243"
date: "November 30, 2016"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Display Functionality}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Print and Summary Functions
The print and summary functions provided by the `mfa` class give you decent summary of the characteristics of your analysis. We can set up using the example dataset:
```{r}
data = read.csv("https://raw.githubusercontent.com/ucb-stat243/stat243-fall-2016/master/problem-sets/final-project/data/wines.csv")
# The first column is just wine IDs, and the last few columns are supplementary physical data.
# Columns 2-54 are the actual sets of ratings.
analysis_data = data[,2:54]
variable_tables = list(1:6,7:12,13:18,19:23,24:29,30:34,35:38,39:44,45:49,50:53)
library(mfa)
# We pass the data, the table specification, the number of components, and arguments to the `scale`
# function in order to z-score the data.
analysis_result = mfa(analysis_data, variable_tables, ncomps = 2, center = TRUE, scale = TRUE)
```

### Summary
The `Summary.R` package provides a single function, `sumeigen`. Simply call it on the `mfa` object:
```{r}
sumeigen(analysis_result)
```

Here you can see statistics about the top-2 latent factors. The eigenvalue, $\lambda$ is the related to the covariance explained of the compromise table. The inertia, $\tau$, is just the percentage of the sum of all eigenvalues this particular eigenvalue is. This is a more interpretable version of $\lambda$.

### Print
The `print.R` package provides several print functions, the most important of which is simply `print`. `print` prints some information about the whole analysis, and also generates some per-table plots. Which table the plots are generated for is selected using the `tablenumber` argument.
```{r}
print(analysis_result, tablenumber=3)
```

As you can see, the function outputs the number of observations, tables, and components. It also plots the observations in terms of the compromise factors, and the partial factor scores and variable loadings for each table. The partial factor scores are simply the projections of a given table's variables into the compromise space for every observation. It is essentially saying *given this compromise space*, what would the points be if this table contained the only variables? The variable loadings for a table just display how much each variable in the table contributes to each componenet.

### Shiny App
To see these visualizations using a web menu, you can run the example shiny app:
```{r, eval=FALSE}
runExample()
```
