---
title: "Using the MFA package"
author: "Stat 243"
date: "11/24/2016"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the MFA package}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
## Package Installation
```{r install,echo=FALSE}
devtools::install_git("https://github.com/cmcneil/project-243","mfa")
```
## Example - Wines dataset
To demonstrate an example of how to use the package, we will use the wines dataset. This dataset consists of ratings for 12 different wines from 10 different wine critics. The critics were forced to rate the wines according to several standardized rating scales, as well as add their own ratings according to whatever rating system they wished. See [Abdi et al, 2013](https://www.utdallas.edu/~herve/abdi-WiresCS-mfa-2013.pdf)

We can get the dataset as follows:
```{r}
# Download the dataset
data = read.csv("https://raw.githubusercontent.com/ucb-stat243/stat243-fall-2016/master/problem-sets/final-project/data/wines.csv")
# The first column is just wine IDs, and the last few columns are supplementary physical data.
# Columns 2-54 are the actual sets of ratings.
data1 = data[,2:54]
```
The `mfa` class requires a `sets` argument, specifying which columns belong to each table respectively. `sets` is a list of vectors specifying the column indices. The following `sets` argument shows the columns for each table.
```{r]}
sets = list(1:6,7:12,13:18,19:23,24:29,30:34,35:38,39:44,45:49,50:53)
```

To use, the package, simply instantiate an `mfa` object, which will do the computations necessary for the core analysis, and then access the various results of the analysis using the object's methods and slots:
```{r}
library(mfa)
result = mfa(data1, sets, ncomps = 2, center = TRUE, scale = TRUE)
```